### 2021-12-03 paper/output/document/application.tex
will typos

%
\begin{isabellebody}%
\setisabellecontext{application}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Applications \label{sec:applications}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this chapter, I 
demonstrate my system's ability to formalize longer, more complicated ethical arguments and present 
the additional capabilities necessary to use my system in practice. In Chapter 2
and Chapter 3, I performed metaethical reasoning while testing different formalizations of the FUL. Metaethical
reasoning analyzes properties of moral thought itself and involves questions about the nature of ethical
truth. This kind of reasoning contrasts with ``applied 
ethical reasoning," which is the use of ethics to resolve dilemmas and make judgements about 
what an agent should or should not do. Applied ethical reasoning is relevant with respect to an agent's
particular situation, not just to an ethical theory as an abstract entity. In previous chapters' application tests,
I performed some toy examples of applied ethical reasoning, but in this chapter, I expore how my system can 
perform extended applied ethical reasoning. Metaethical reasoning is most interesting to philosophers 
who are trying to formulate the ``best" theory of ethics. Applied ethical reasoning, on the other hand, 
is useful to ordinary people who are trying to decide how to live their lives. In order for automated 
ethics to guide AI agents, it must perform applied ethical reasoning.

@@

The specific kind of common sense reasoning required appears to be a challenge 
for automating Kantian ethics, and perhaps a reason that consequentialist or virtue ethical automated
agents would be within closer reach for automation. Ultimately, Kantian ethics is still
easiest to automate because it will require fewer, less controversial common sense facts than other ethical
systems. As the examples in this section demonstrate, Kantian ethics requires a definition of lying (which any other 
theory would also requires) and the knowledge that if everyone lies in a given context, no one will believe 
each other in that particular context. This latter fact may not be required for every ethical theory, 
but is relatively uncontroversial. It is a kind of intuition about human behavior that is generally accepted. 
Neither a definition of lying nor this property of lying seem like unreasonable prerequisites for ethical reasoning.

@@

This chapter will provide additional examples of the kinds of common sense facts and maxim formulation required to get my system
off the ground. I will aim to use as lean and uncontroversial of a common sense database as possible
to achieve robust and powerful results. This serves as evidence for the ease of automating
Kantian ethics, an example of what additional work my system requires, and a demonstration of the contributions
that I make. These examples demonstrate that nuanced common sense facts and maxims can cause my system to contribute 
nuanced judgements faithful to philosophical literature. On one hand, this means that my system can perform
sophisticated ethical reasoning, but on the other, the quality of this reasoning is heavily reliant on 
the common sense database and the input maxim. Thus, my system \emph{could} make smart ethical judgements 
in practice, but getting to that point will require a robust, trusted common sense database and maxim
formulator.

@@

    (a) the act requires knowingly uttering a falsehood 

@@

While it is true that lying to the murderer should be permissible, Korsgaard notes that many will want
to say something stronger, like the fact that lying to the murderer is obligatory in order to protect
the intended victim \citep[15]{KorsgaardRTL}. It seems like we would be doing something wrong if I revealed the victim's
location, knowing that this revelation would cost them their life. Korsgaard solves this problem by 
noting that, while the FUL shows that lying to the murderer permissible, other parts of Kant's ethics
show that it is obligatory. Recall that Kant presents perfect and imperfect duties,
where the former are strict, inviolable, and specific and the latter are broader prescriptions for action.
Perfect duties always supersede imperfect duties when the two conflict.
For example, the duty to not murder is a perfect duty and the duty to give to charity is an imperfect duty. 
The FUL generates perfect duties and Kant's extended theory of virtues generates imperfect duties. The details
of this theory and these distinctions are outside the scope of this paper, but the crucial note is that 
other parts of Kant's ethical theory generate the obligation to lie to the murderer. I chose
to formalize the FUL because it is, in some sense, the strongest of the formulations of the categorical
imperative. An even more sophisticated Kantian reasoner could formalize his theory of virtue and his
other formulatations of the categorical imperative in order to generate the obligation to lie to the murderer, 
but the FUL is the strongest and most foundational of these principles. The fact that my system merely
shows that lying to the murderer is permissible, but not obligatory is consistent with the part of 
Kant's ethical theory that I formalize and demonstrates that I have faithfully implemented the FUL.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
While this example demonstrates the power of my system (when equipped with some common sense), it 
also shows how vital the role of the common sense reasoning is. Slight, intuitive changes in the common
sense facts achieved totally different conclusions about lying. This represents an obstacle to 
fully automated ethical reasoning; such an agent would need a trusted database of common 
sense facts, which is still an unsolved problem. My work is one step towards such an agent, but the 
importance of common sense means that much progress must be made in order to completely automate ethics.

@@

The Kantian response to this criticism is to require that the circumstances included in the formulation
of the maxim be ``morally relevant." In the example above, my purple shirt and the date clearly have no bearing on 
the moral status of lying. On the other hand, consider the maxim, ``When I am unmeployed, I will murder
someone in order to take their job." The circumstances of being unmeployed clearly have some bearing on the moral
relevance of the murder in question; they speak to the motivation for the murder. While this view seems 
to track how we actually perform moral reasoning, it leaves open the question of how to determine
which circumstances are morally relevant. Here, O'Niell reminds us that the Formula of Universal Law is 
a ``test of moral worth rather than of outward rightness" \citep[98]{constofreason}. The FUL is a way 
for an agent to decide how they should behave, not for a third-party to judge their behavior. Ethics is 
a personal process for Kant, so the FUL is designed to help agents internally make decisions, not to 
judge others' decisions. Because agents use the FUL to evaluate their own behavior, the test is at its 
best when they make a good faith effort to isolate the \emph{principle} of their action, rather than some
``surface intent" \citep[87]{constofreason}. The FUL is supposed to determine if an agent's principle of action
is universally consistent, so it is at its most effective when an agent accurately formulates the principle
they act on. Circumstances are morally relevant if they accurately reflect the way that the agent is 
thinking about their own action. In the example above, the circumstances of wearing a purple shirt don't reflect
the principle of the liar's action. Their inclusion is clearly a disingenous attempt to evade the universalizability
test, but because the FUL is a test of personal integrity, it cannot withstand this kind of mental
gymnastics.

>>>>

%
\begin{isabellebody}%
\setisabellecontext{application}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Applications \label{sec:applications}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this chapter, I 
demonstrate my system's ability to formalize longer, more complicated ethical arguments and present 
the additional capabilities necessary to use my system in practice. In Chapter 2
and Chapter 3, I performed metaethical reasoning while testing different formalizations of the FUL. Metaethical
reasoning analyzes properties of moral thought itself and involves questions about the nature of ethical
truth. This kind of reasoning contrasts with ``applied 
ethical reasoning," which is the use of ethics to resolve dilemmas and make judgements about 
what an agent should or should not do. Applied ethical reasoning is relevant with respect to an agent's
particular situation, not just to an ethical theory as an abstract entity. In previous chapters' application tests,
I performed some toy examples of applied ethical reasoning, but this chapter is an extended exploration of
my system's ability to perform applied ethical reasoning. Metaethical reasoning is most interesting to philosophers 
who are trying to formulate the ``best" theory of ethics. Applied ethical reasoning, on the other hand, 
is useful to ordinary people who are trying to decide how to live their lives. In order for automated 
ethics to guide AI agents, it must perform applied ethical reasoning.

@@

The specific kind of common sense reasoning required appears to be a challenge 
for automating Kantian ethics, and may imply that consequentialist or virtue ethical automated
agents are within closer reach for automation. Ultimately, Kantian ethics is still
easiest to automate because it will require fewer, less controversial common sense facts than other ethical
systems. As the examples in this section demonstrate, Kantian ethics requires a definition of lying (which any other 
theory would also requires) and the knowledge that if everyone lies in a given context, no one will believe 
each other in that particular context. This latter fact may not be required for every ethical theory, 
but is relatively uncontroversial. It is a kind of intuition about human behavior that is generally accepted. 
Neither a definition of lying nor this property of lying seem like unreasonable prerequisites for ethical reasoning.

@@

This chapter will provide additional examples of the kinds of common sense facts and maxim formulation required to get my system
off the ground. I will aim to use a lean and uncontroversial common sense database
to achieve robust and powerful results. This serves as evidence for the ease of automating
Kantian ethics, an example of what additional work my system requires, and a demonstration of the contributions
that I make. These examples demonstrate that nuanced common sense facts and maxims can cause my system to contribute 
nuanced judgements faithful to philosophical literature. On one hand, this means that my system can perform
sophisticated ethical reasoning, but on the other, the quality of this reasoning is heavily reliant on 
the common sense database and the input maxim. Thus, my system \emph{could} make smart ethical judgements 
in practice, but getting to that point will require a robust, trusted common sense database and maxim
formulator.

@@

    (a) the act requires knowingly uttering a falsehood and,

@@

While it is true that lying to the murderer should be permissible, Korsgaard notes that many will want
to say something stronger, like the fact that lying to the murderer is obligatory in order to protect
the intended victim \citep[15]{KorsgaardRTL}. It seems like we would be doing something wrong if I revealed the victim's
location, knowing that this revelation would cost them their life. Korsgaard solves this problem by 
noting that, while the FUL shows that lying to the murderer permissible, other parts of Kant's ethics
show that it is obligatory. Recall that Kant presents perfect and imperfect duties,
where the former are strict, inviolable, and specific and the latter are broader prescriptions for action.
Perfect duties always supersede imperfect duties when the two conflict.
For example, the duty to not murder is a perfect duty and the duty to give to charity is an imperfect duty. 
The FUL generates perfect duties and Kant's extended theory of virtues generates imperfect duties. The details
of this theory and these distinctions are outside the scope of this paper, but the crucial note is that 
other parts of Kant's ethical theory generate the obligation to lie to the murderer. I chose
to formalize the FUL because it is, in some sense, the strongest of version of the categorical
imperative. An even more sophisticated Kantian reasoner could formalize his theory of virtue and his
other formulatations of the categorical imperative in order to generate the obligation to lie to the murderer, 
but the FUL is the strongest and most foundational of these principles. The fact that my system merely
shows that lying to the murderer is permissible, but not obligatory is consistent with the part of 
Kant's ethical theory that I formalize and demonstrates that I have faithfully implemented the FUL.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
While this example demonstrates the power of my system (when equipped with some common sense), it 
also shows how vital the role of the common sense reasoning is. Slight, intuitive changes in the common
sense facts achieved totally different conclusions about lying. This represents an obstacle to 
fully automated ethical reasoning; such an agent would need a trusted database of common 
sense facts, which is still an unsolved problem. My work is one step towards such an agent, but the 
importance of common sense means that much progress must be made in order to completely automate ethics.

@@

The Kantian response to this criticism is to require that the circumstances included in the formulation
of the maxim be ``morally relevant." In the example above, my purple shirt and the date clearly have no bearing on 
the moral status of lying. On the other hand, consider the maxim, ``When I am unemployed, I will murder
someone in order to take their job." The circumstances of being unemployed clearly have some bearing on the moral
relevance of the murder in question; they speak to the motivation for the murder. While this view seems 
to track how we actually perform moral reasoning, it leaves open the question of how to determine
which circumstances are morally relevant. Here, O'Niell reminds us that the Formula of Universal Law is 
a ``test of moral worth rather than of outward rightness" \citep[98]{constofreason}. The FUL is a way 
for an agent to decide how they should behave, not for a third-party to judge their behavior. Ethics is 
a personal process for Kant, so the FUL is designed to help agents internally make decisions, not to 
judge others' decisions. Because agents use the FUL to evaluate their own behavior, the test is at its 
best when they make a good faith effort to isolate the \emph{principle} of their action, rather than some
``surface intent" \citep[87]{constofreason}. The FUL is supposed to determine if an agent's principle of action
is universally consistent, so it is at its most effective when an agent accurately formulates the principle
they act on. Circumstances are morally relevant if they accurately reflect the way that the agent is 
thinking about their own action. In the example above, the circumstance of wearing a purple shirt doesn't reflect
the principle of the liar's action. Its inclusion is clearly a disingenous attempt to evade the universalizability
test, but because the FUL is a test of personal integrity, it cannot withstand this kind of mental
gymnastics.
